---
title: "Leaf-level coordination principles propagate to the ecosystem scale"
author: "Ulisse Gomarasca et al."
date: "2023-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Description

This R Markdown document guides you through the code used for the analysis in the
manuscript titled: "Leaf-level coordination principles propagate to the ecosystem scale"
by Gomarasca et al., submitted at Nature Communications.

For the analysis, we used R version 4.1.0 (2021-05-18) "Camp Pontanezen", and
RStudio 2022.12.0+353 "Elsbeth Geranium" Release (7d165dcfc1b6d300eb247738db2c7076234f6ef0, 2022-12-03) for Windows.
The main packages used are described in the Methods section of the manuscript.




## Manuscript analysis and figures

Here, we run the code used to generate the principal component analysis output 
(PCA_stats_Figure_x) and plots for Figures 1, 2, and 3. The input data subsets
to the PCAs are also generated.
Supplementary Tables 2, 3, and 4 are compiled from a combination of the output
statistics in this chunk and the subsequent chunks.
Additionally, here we generate a list of the sites used for all the analyses, and
Supplementary Table 7 with the relative number of sites for each IGBP vegetation
class in each analysis.

The running time for the code can take a few hours with the full 499 bootstrapping
iterations.
All the output files can be found in the "results" folder of this R project.

```{r}
rm(list = ls(all = TRUE)) # clear environment

source("scripts/functions/PCAs.R") # function to run the analyses
source("scripts/functions/plot_PCAs.R") # function to plot analyses' results

print("Running PCA analysis to reproduce results of the manuscript titled: 'Leaf-level coordination principles propagate to the ecosystem scale'. This can take a few hours.")

PCAs(figs = 1:3, boots = 499)  # run Principal Component Analysis: 1 for Figure 1;
# 2 for Figure 2; 3 for Figure 3.
# Lower the 'boots' argument to run the analysis faster (but consider that the
# results will not match the ones in the manuscript if you do).

plot_PCAs(figs = 1:3) # plot PCA output: 1 for Figure 1; 2 for Figure 2; 3 for Figure 3.
```


## Supplementary Figures 1, 3-9

Here, we run the principal component analysis for the Least-Cost Hypothesis at
the ecosystem scale with different variables, and plot the output into
Supplementary Figures 4 to 7.
We also run the PCA with i) all sites, ii) forest sites, and iii) evergreen
needle-leaf forest sites, and plot the output into Supplementary Figures 1, 3,
and 8.
We generate the same output and input data table structure as for the main figures,
which can be found in the 'results' folder.

```{r}
rm(list = ls(all = TRUE)) # clear environment

source("scripts/functions/PCAs.R") # function to run the analyses
source("scripts/functions/plot_PCAs.R") # function to plot analyses' results

print("Running PCA analysis to reproduce Supplementary Figures 1, 3, 4, 5, 6, 7, and 8 in the Supplementary Information for the manuscript titled: 'Leaf-level coordination principles propagate to the ecosystem scale'. This can take a few hours.")

PCAs(figs = c(11:13, 21, 22, 31, 32, 35:38), boots = 499) # run principal component analysis to generate results for the supplementary figures:
# 11:12 for Supplementary Figure 1;
# 13 for Supplementary Figure 3;
# 21:22 for Supplementary Figure 4;
# 31:32 for Supplementary Figure 9;
# 35:38 for Supplementary Figures 5-8;
# 101 for additional analyses related to the leaf economics spectrum at the ecosystem scale (not included in the supplementary information)

plot_PCAs(figs = c(1, 11:13, 2, 21, 22, 3, 31, 32, 35:38), plot_main = F) # plot PCA results for the selected supplementary figures.
# For comparison, we re-plot the results from the main analyses together with all sites together with 'all forests' and 'evergreen needle-leaf forests' (Supplementary Figures 1, 3, 8), but we avoid re-plotting the main figures with the 'plot_main' argument set to false. Make sure to run the 'Manuscript analysis and figures' chunk if the results of the main analysis are not saved in the 'results' folder yet.
```


## Supplementary Figure 2

Here, we run a set of predictive analysis that include multimodel inference and
relative importance to plot Supplementary Figure 2.

```{r}
rm(list = ls(all = TRUE)) # clear environment

print("Running multimodel inference and relative importance analysis to reproduce Supplementary Figures 2 in the Supplementary information for the manuscript titled: 'Leaf-level coordination principles propagate to the ecosystem scale'")

source("scripts/multimodel_inference_SuppFig2.R")
```


## Supplementary Table 2

Here, we run a simple linear mixed model to guide the outlook for modeling purposes.

```{r}
rm(list = ls(all = TRUE)) # clear environment

source("scripts/linear_mixed_model_revisionNatComm.R")
```